Handling String-Number Conversion in String Constraint Solving

### 2019-11-15 evaluation.tex
added description of new benchmarks

We have implemented the proposed approach in a prototype tool \textsf{z3-Trau} as a part of the Z3 theorem prover. This approach allows us to efficiently solve not only formulae over string constraints, but also combine string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

We compare \textsf{z3-Trau} with other state-of-the-art string solvers, namely CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, \textsf{z3str3}~\cite{zheng2013z3}, Ostrich~\cite{chen2019decision} and \textsf{Trau+}~\cite{parosh2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints. We present data on two benchmark groups that demonstrate two points. First, only our tool can efficiently solve formulae that contain string-number and number-string conversion constraints and which can be a combination of constraints over multiple theories. Second, we show that our tool is competitive on well-known benchmarks that can be effectively handled by other existing solvers. The summary of experimenting for each benchmark group is given in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. All experiments were executed on a computer with 4-core CPU, 8 GiB RAM and Ubuntu 18.4-LVM OS. The timeout was set to 10s for each test. Columns with heading \texttt{sat}/\texttt{unsat} show numbers of solved formulae. Columns with heading \texttt{timeout} indicate the number of times the solver exceeded the time limit and columns with heading $\times$ show numbers of instances for which the particular solver either reported unknown or reported incorrect answer or crashed.

@@

\hide{
In this section, we compare our implementation z3-Trau with other SMT tools cvc4, z3, and z3-str3 as evauation. To show the general performance of z3-Trau, we compare z3-Trau with other string solvers on selected string benchmarks: PyEx is a benchmark obtained from symbolic execution of Python code[]; APLAS is a benchmark involving looping word equations[]; 
LeetCode is obtiained from concolic testing LeetCode solutions written in Python code; StringFuzz is a benchmark of instance SMT-lib string problems generated by StringFuzz generator tool[]; cvc4\textsubscript{pred} and cvc4\textsubscript{term} are benchmarks provided by the cvc4 development team[]. Table~\ref{table:base_benchmark} shows the result of the comparison. The experiments are conducted with machines of the following specifications: 4-core CPU, 8GB RAM, Ubuntu 18.4-LVM OS. We set the timeout is to 10 seconds. Because the amount of problems is very large, we ran these experiments separately on several machines with the same specification on a computer cluster. The results are either sat, unsat, timeout, or $\times$. In case $\times$, the result may be unknown, error, or exception.
}

@@

\textbf{Comparison according to Table 2.....}

>>>>

We have implemented the proposed approach in a prototype tool \tool as a part of the Z3 theorem prover. This approach allows us to efficiently solve not only formulae over string constraints, but also combine string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

We compare \tool with other state-of-the-art string solvers, namely CVC4~\cite{cvc4Tool}, Z3~\cite{z3}, \textsf{z3str3}~\cite{zheng2013z3}, Ostrich~\cite{chen2019decision} and \textsf{Trau+}~\cite{parosh2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints. We present data on two benchmark groups that demonstrate two points. First, only our tool can efficiently solve formulae that contain string-number and number-string conversion constraints and which can be a combination of constraints over multiple theories. Second, we show that our tool is competitive on conversion-less benchmarks as well. The summary of experimenting for each benchmark group is given in Table~\ref{table:base_benchmark} and Table~\ref{table:str_int_benchmark}, respectively. All experiments were executed on a computer with 4-core CPU, 8 GiB RAM and Ubuntu 18.4-LVM OS. The timeout was set to 10s for each test. Columns with heading \texttt{sat}/\texttt{unsat} show numbers of solved formulae. Columns with heading \texttt{timeout} indicate the number of times the solver exceeded the time limit and columns with heading $\times$ show numbers of instances for which the particular solver either reported unknown or reported incorrect answer or crashed.

@@

\paragraph{A Comparison on benchmarks containing string-number and number-string constrains.}
Our second group of benchmarks was created in order that we could compare our proposed approach for solving string-number and number-string conversions with existing approaches. This group contains a total of 3 benchmarks: full\_str\_int, filtered\_str\_int and rec\_fun. None of these benchmarks were artificially generated but were created from real Python's and javascript's codes.

The first benchmark in the group is full\_str\_int, which is a collection of SMT queries. This collection was generated by applying a tool for concolic testing to Python codes selected from the previously mentioned LeetCode platform and from Python core libraries. All selected Python codes use the \texttt{int()} function, which converts a string into a number system based on the specified base. The Benchmark consists of 21573 tests in total.

The second benchmark, called filtered\_str\_int, is a subset of the previous full\_str\_int benchmark. The filtered\_str\_int benchmark was created by removing tests where cvc4 reported UNSAT and where cvc4's returned unsat core contained no conversion constraint. This benchmark was created in order that we better compare individual conversion approaches. A total of 7396 tests were left.

The last benchmark in the group is rec\_fun, which is a collection of javascript functions that were handcrafted encoded into smt2 format using recursive functions. Besides running examples from the introduction, the benchmark also includes \texttt{split} and \texttt{replaceAll} functions. In total, we managed to create 43 tests that combine several SMT theory and contain conversion constrains.

\hide{
In this section, we compare our implementation z3-Trau with other SMT tools cvc4, z3, and z3-str3 as evauation. To show the general performance of z3-Trau, we compare z3-Trau with other string solvers on selected string benchmarks: PyEx is a benchmark obtained from symbolic execution of Python code[]; APLAS is a benchmark involving looping word equations[]; 
LeetCode is obtiained from concolic testing LeetCode solutions written in Python code; StringFuzz is a benchmark of instance SMT-lib string problems generated by StringFuzz generator tool[]; cvc4\textsubscript{pred} and cvc4\textsubscript{term} are benchmarks provided by the cvc4 development team[]. Table~\ref{table:base_benchmark} shows the result of the comparison. The experiments are conducted with machines of the following specifications: 4-core CPU, 8GB RAM, Ubuntu 18.4-LVM OS. We set the timeout is to 10 seconds. Because the amount of problems is very large, we ran these experiments separately on several machines with the same specification on a computer cluster. The results are either sat, unsat, timeout, or $\times$. In case $\times$, the result may be unknown, error, or exception.

@@

\textbf{Comparison according to Table 2.....}
}
